Reproducible development environments with rix

### 2023-12-21 paper.qmd
nixpkgs

For example, the `{xlsx}` package requires the Java programming language to be
installed on your computer to successfully install. This can be difficult to
achieve, and `{xlsx}` bullied many R developers throughout the years (especially
those using a Linux distribution, `sudo R CMD javareconf` still plagues my
nightmares). But with Nix, it suffices to declare that we want the `{xlsx}`
package for our project, and Nix figures out automatically that Java is required
and installs and configures it. It all just happens without any required
intervention from the user. The second advantage of Nix is that it is possible
to pin a certain revision of the Nix packages’ repository (called nixpkgs) for
our project. Pinning a revision ensures that every package that Nix installs
will always be at exactly the same versions, regardless of when in the future
the packages get installed.

>>>>

For example, the [xlsx]{.pkg} package requires the [Java]{.proglang} programming
language to be installed on your computer to successfully install. Depending on
the system one tries to install [xlsx]{.pkg}, installing [Java]{.proglang} might
not easy, nor even possible. But with [Nix]{.proglang}, it suffices to declare
that the [xlsx]{.pkg} package is needed for the project, and [Nix]{.proglang}
figures out automatically that [Java]{.proglang} is required and installs and
configures it. It all just happens without any required intervention from the
user.

But where do these packages come from? When installing a package using Nix an
expression written in the Nix programming language gets downloaded from the
`nixpkgs` Github repository and evaluated. The evaluation of these expressions
results in the installation of the package, alongside its dependencies. Because
`nixpkgs` is a collection of expressions hosted on Github, it is possible to pin
a specific revision of `nixpkgs` to ensure reproducibility of our project.
Pinning a revision ensures that every package that Nix installs will always be
at exactly the same versions, regardless of when in the future the packages get
installed.
